A user-friendly tool to evaluate the effectiveness of no-take marine reserves

### 2017-12-15 Manuscript.tex
updated manuscript with comments from Melaina

\documentclass[12pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={A user--friendly tool to evaluate the effectiveness of no--take marine reserves},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\setlength{\droptitle}{-2em}
  \title{A user--friendly tool to evaluate the effectiveness of no--take marine
reserves}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
\subtitle{Marine reserve evaluation}
  \author{}
  \preauthor{}\postauthor{}
  \date{}
  \predate{}\postdate{}

@@

Marine reserves are implemented to achieve a variety of objectives, but
are seldom rigorously evaluated to determine whether those objectives
are met. In the rare cases when evaluations do take place, they
typically focus on ecological indicators and ignore other relevant
objectives such as socioeconomics and governance. And regardless of the
objectives, the diversity of locations, monitoring protocols, and
analysis approaches hinder the ability to compare results across case
studies. Moreover, analysis and evaluation of reserves is generally
conducted by outside researchers, not the reserve managers or users,
plausibly thereby hindering effective local management and rapid
response to change. We present a framework and tool, called ``MAREA'',
to overcome these challenges. Its purpose is to evaluate the extent to
which any given reserve has achieved its stated objectives. MAREA
provides specific guidance on data collection and formatting, and then
conducts rigorous causal inference analysis based on data input by the
user, providing real--time outputs about the effectiveness of the
reserve. MAREA's ease of use, standardization of state--of--the--art
inference methods, and ability to analyze marine reserve effectiveness
across ecological, socioeconomic, and governance objectives could
dramatically further our understanding and support of effective marine
reserve management.

@@

Empirical evidence shows that MPAs increase biomass {[}4,7{]}, enhance
resilience to climatic impacts {[}8,9{]}, and preserve genetic diversity
{[}10{]}. Compared to MPAs that grant partial protection, marine
reserves have higher levels of biomass, density, richness, and larger
organisms {[}3,11--13{]}. However, these effects are often measured as
biological changes within the reserves through time, and many lack a
control site for comparison {[}14{]}. This approach does not account for
other factors (\emph{e.g.} system-level changes in productivity caused
by predatory release {[}15{]}; or favorable environmental conditions
{[}16{]}) for which one must control {[}17{]} in order to causally
attribute a biological change to the reserve. Other studies have used a
control--impact comparison approach that uses control sites but does not
address temporal variability {[}4,7,18--20{]}.

A smaller fraction of studies have used a before--after--control--impact
(\emph{i.e.} BACI) design comparing reserves to control sites before and
after implementation {[}4,21,22{]}, which allows the use of causal
inference techniques that estimate the effect of the reserve. For
example, in ref {[}21{]} authors use a BACI design and observe increases
in lobster catches --a proxy for abundances-- after the implementation
of the reserves for protected and control sites. However, the temporal
changes in the reserve were greater than in the control site, suggesting
a positive effect off the reserve on lobster catches. But even when
proper causal inference can be drawn, results are often different across
reserves. Effects of reserves on ecological and economic outcomes are
highly heterogeneous, and often depend on the specific ecological,
economic, and social context.

@@

An increasingly popular way to make science accessible, reproducible,
scalable, and replicable is through Open Science and the development of
open--access tools {[}28{]}. The Ocean Health Index {[}29,30{]}, for
example, successfully standardized a way to measure the health and
benefits of the oceans. This approach has been implemented at a global
scale, but also at country-level {[}31{]}, and regionally {[}32,33{]}.
Open access tools are not limited to conservation, and have also been
developed to evaluate fishery performance {[}34,35{]}, design
territorial use rights for fisheries {[}36{]}, and improve decision
making in the hydro power industry {[}37{]}.

The purpose of this paper is to describe a user--friendly tool, called
``MAREA'', to rigorously systematize the evaluation of marine reserve
effectiveness in terms of fisheries and marine conservation. The tool is
in the form of an open--source application that uses state--of--the--art
methods from program evaluation to compare a reserve to control sites
along a number of biological, economic, and governance dimensions. We
first provide a list of commonly stated management objectives and match
them to appropriate indicators. We then develop a simple approach to
analyzing these indicators building on causal inference techniques
{[}21{]}, which help us understand the effect of management
interventions {[}25,38{]}. To implement the analytical approach, we
introduce the Marine Reserve Evaluation Application (MAREA), an open
source, web--based tool that automates the framework described in this
paper and enables its broader use. Finally, we present a case study on
the evaluation of a marine reserve established by the fishers of Isla
Natividad (Mexico) in 2006, to demonstrate the potential of MAREA.

@@

Throughout this study, we will refer to the stated goals for which a
marine reserve was designed as ``objectives.'' This work was motivated
by the case of Mexico, where 39 reserves have been implemented over the
past five years to achieve objectives such as increasing productivity in
nearby waters or recovery of overexploited species; most of these
reserves have never been formally evaluated for effectiveness at meeting
those objectives. Thus, our focus was on identifying common objectives
of marine reserves in Mexico. However, a literature review and
discussions with marine reserve researchers suggest that the objectives
driving Mexican marine reserve implementation are similar to those in
the rest of the world. Thus, we group these objectives into seven major
categories that may be applied to marine reserves worldwide. Any given
reserve may have been implemented to meet one or more of these. The list
includes objectives stated in legislation {[}39,40{]} and official
documents such as the Technical Justification Studies (\emph{Estudios
Técnicos Justificativos}), agreements, and decrees associated to these
areas:

@@

Based on these seven objectives, we determined a set of associated
indicators to evaluate reserve effectiveness. These indicators are
specific variables on which data could be collected, and analyzed, to
ultimately determine whether the corresponding objective was causally
being achieved by the marine reserve. The list of indicators was
compiled through a review of scientific literature in which we
identified indicators that were used to measure similar
objectives{[}3--5,7,11,13,14,18--21,23,24,41--44{]}. A first filter
eliminated indicators for which baseline data do not typically exist in
Mexico. The preliminary list of indicators was reviewed at a workshop
with participation of members from Mexican fishery management agencies
and non-government organizations. Later, these were presented to fishers
from the Ensenada Fishing Cooperative (\emph{S.C.P.P. Ensenada}), in El
Rosario, Baja California, who provided input. Our final list of
indicators includes those identified in review works {[}4,44{]}.

Indicators are divided into three main categories: biological,
socioeconomic, and governance (Table 1). The nine biological indicators
focus on fish and invertebrate communities that are evaluated using
underwater ecological surveys performed inside and outside the reserve
(see Data and Analysis section for specific sampling design and
methodologies). Five socioeconomic indicators reflect the performance of
the fishery in terms of landings, income from landings, and availability
of alternative livelihoods. Fifteen governance indicators describe the
governance structures under which the community operates (\emph{e.g.},
access rights to the fishery, number of fishers, legal recognition of
the reserve). Most biological and socioeconomic indicators are
quantitative and require a numerical entry (\emph{e.g.} Fish biomass)
while all governance indicators, one biological indicator, and one
socioeconomic indicator are qualitative and rely on a descriptive entry
(\emph{e.g.} Reasoning for reserve location). Many of them specifically
measure an outcome of the reserve, though some are designed to further
the understanding of the mechanisms driving a reserve's performance. In
that sense, most biological and socioeconomic indicators are outcome
variables. On the other hand, governance indicators are viewed as
possible explanatory variables of reserve performance. Whenever an
indicator is applied to ``Target species'', it means that the indicator
can be used for all species (\emph{e.g.} Fish Biomass) and/or for
individual species that are either the conservation target of the
reserve or are of particular economic or ecological interest
(\emph{e.g.} Grouper Biomass). Finally, indicators B3 and B4 are
different in that B3 only looks at the density of organisms above size
at first maturity (related to reproductive potential), while B4 measures
the density of all fish or of a target species. Each indicator targets
different plausible desired outcomes, like increased reproductive
potential (\emph{i.e.} B3; {[}45{]}) or having more fish -regardless of
their size- to attract tourism (\emph{i.e.} B4). Table 1 presents the
proposed indicators, and Table 2 shows how objectives are matched with
biological and socioeconomic indicators. Governance indicators are
excluded from Table 2, as governance indicators are important to
consider for every objective.

\begin{table}

\caption{\label{tab:unnamed-chunk-2}List of indicators to evaluate the effectiveness of no-take marine reserves.}
\centering
\resizebox{\linewidth}{!}{\begin{tabular}[t]{l|l|l|l}
\hline
Code & Indicator & Data type & Unit\\
\hline
\multicolumn{4}{l}{\textbf{Biological}}\\
\hline
\hspace{1em}B1 & Shannon diversity index & Continuous & \\
\hline
\hspace{1em}B2 & Species richness & Discrete & Number of species/transect\\
\hline
\hspace{1em}B3 & Density of mature organisms & Continuous & Percent\\
\hline
\hspace{1em}B4 & Density* & Continuous & Organisms/transect\\
\hline
\hspace{1em}B5 & Natural Disturbance & Descriptive & \\
\hline
\hspace{1em}B6 & Mean Trophic Level & Continuous & \\
\hline
\hspace{1em}B7 & Biomass* & Continuous & kg/transect\\
\hline
\multicolumn{4}{l}{\textbf{Socioeconomic}}\\
\hline
\hspace{1em}S1 & Total landings* & Continuous & kg\\
\hline
\hspace{1em}S2 & Income from total landings* & Continuous & \$\\
\hline
\hspace{1em}S3 & Alternative economic opportunities & Ordinal & \\
\hline
\multicolumn{4}{l}{\textbf{Governance}}\\
\hline
\hspace{1em}G1 & Access to the fishery & Categorical & \\
\hline
\hspace{1em}G2 & Number of fishers & Discrete & \\
\hline
\hspace{1em}G3 & Legal recognition of reserve & Binary & \\
\hline
\hspace{1em}G4 & Reserve type & Descriptive & \\
\hline
\hspace{1em}G5 & Illegal harvesting & Ordinal & \\
\hline
\hspace{1em}G6 & Management plan & Binary & \\
\hline
\hspace{1em}G7 & Reserve enforcement & Descriptive & \\
\hline
\hspace{1em}G8 & Size of reserve & Discrete & \\
\hline
\hspace{1em}G9 & Reasoning for reserve location & Descriptive & \\
\hline
\hspace{1em}G10 & Membership to fisher organizations & Binary & \\
\hline
\hspace{1em}G11 & Type of fisheries organizations & Categorical & \\
\hline
\hspace{1em}G12 & Representation & Ordinal & \\
\hline
\hspace{1em}G13 & Internal Regulation & Binary & \\
\hline
\hspace{1em}G14 & Perceived Effectiveness & Categorical & \\
\hline
\hspace{1em}G15 & Social Impact of Reserve & Categorical & \\
\hline
\multicolumn{4}{l}{\textsuperscript{*} The indicator is also applied to target species}\\
\end{tabular}}
\end{table}

@@

\begin{table}

\caption{\label{tab:unnamed-chunk-3}Management objectives and respective performance indicators.}
\centering
\resizebox{\linewidth}{!}{\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}|l|l|l|l|l|l|l|l|l|l|l|l|l|l}
\hline
Objective & B1 & B2 & B3 & B4 & B4* & B5 & B6 & B7 & B7* & S1 & S1* & S2 & S2* & S3\\
\hline
Avoid overexploitation &  &  & x & x & x & x & x & x & x & x & x & x & x & x\\
\hline
Conserve species under a special protection &  &  & x &  & x & x &  &  & x & x &  & x &  & x\\
\hline
Maintain biological process & x & x &  & x &  & x & x & x &  &  &  &  &  & x\\
\hline
Improve fishery production in nearby waters &  &  &  & x & x & x &  & x & x & x & x & x & x & x\\
\hline
Preserve biological diversity and the ecosystem & x & x &  & x &  & x & x & x &  &  &  &  &  & x\\
\hline
Recover overexploited species &  &  & x &  & x & x &  &  & x &  & x &  & x & x\\
\hline
Recover species of economic interest &  &  & x &  & x & x &  &  & x &  & x &  & x & x\\
\hline
\multicolumn{15}{l}{\textsuperscript{*} The indicator is applied to target species}\\
\end{tabular}}
\end{table}

@@

Socioeconomic data are often collected by fishers, natural resource
management agencies, or CSOs by recording landings, income, and
sometimes prices for each species. To control for inflation, income is
adjusted with the country's consumer price index {[}48{]}:

@@

While this model does not formally allow for causal inference, we can
still measure changes in mean landings and income before and after the
implementation of the reserve and provide valuable input. For both
models (eq. 1 and eq. 3), we estimate the model coefficient with
ordinary least squares, and calculate heteroskedastic--robust standard
errors {[}49{]}.

@@

We developed MAREAin R version 3.4.2 and R Studio 1.1.383 {[}51{]} using
the Shiny package {[}52{]}, to build an interactive web application
hosted on an open server; the MAREA app can be accessed at
\href{turfeffect.shinyapps.io/marea/}{turfeffect.shinyapps.io/marea}.
While the original version was developed in Spanish because it was aimed
for Mexico and other Latin-American countries, all of its content can be
translated by a translation widget available within the app.

MAREA is designed as a 6--step process, divided in tabs that appear upon
launching the app. The first tab introduces the app and summarizes the
evaluation process. Then, the user selects management objectives, which
MAREA automatically matches to appropriate indicators, based on Table 2.
Users can also select and deselect indicators based on their interests
and data availability by ``clicking'' on the check-boxes in MAREA. The
user can then load data on one or more reserves, using standard *.csv
text files; sample datasets are provided within MAREA. Once data have
been loaded, MAREA identifies all reserves in the data, and lets the
user select the reserve to be evaluated. At this point, the user can
also specify the year of implementation of the reserve, reserve
dimensions, and indicate target species that are of particular
management interest. MAREA provides the user with a section to confirm
that all the decisions made leading up to that point are correct. Once
the user has confirmed all input data, objectives, and other
information, MAREA performs the formal program evaluation analyses
discussed above. For a typical data set, the automated analysis step
takes less than one second. Finally, the user is taken to the results
tab where all results are presented in a simple format. The user can
also download a more comprehensive technical report produced in *.pdf
format.

@@

\textbf{Fig 3. General location of Isla Natividad (left) and map of the
island (right).} The marine reserve polygon is indicated in red, and the
approximate location of control sites is indicated by blue squares (B =
Babencho, D = La Dulce). Shapefiles for Mexican coastline and the US
were obtained from the Instituto Nacional de Estadistica, Geografia e
Informatica of Mexico (INEGI:
www.inegi.org.mx/geo/contenidos/geoestadistica/m\_geoestadistico.aspx)
and the tmap R package {[}58{]}, respectively.

In 2006, the Isla Natividad community established a biological baseline
following the data collection protocol described in this study. The
community then implemented two community--based marine reserves within
their TURF {[}8,41,59{]} after establishing a baseline for the
soon--to--be reserves and control sites. Evidence suggest that these
reserves have been effective at enhancing resilience to climate
variations {[}8{]} and preserving genetic diversity of high value
commercial species such as abalone {[}10{]}. These ecological benefits
have been translated into economic benefits, enhancing population
persistence and bolstering abalone fisheries {[}43{]}. For the purpose
of this evaluation, we focused on the ``La Plana / Las Cuevas'' marine
reserve, located at the southern end of the island (Fig. 3) and its
corresponding control site ``La Dulce / Babencho''.

@@

>>>>

\documentclass[12pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={A user-friendly tool to evaluate the effectiveness of no-take marine reserves},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\setlength{\droptitle}{-2em}
  \title{A user-friendly tool to evaluate the effectiveness of no-take marine
reserves}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
\subtitle{Marine reserve evaluation}
  \author{}
  \preauthor{}\postauthor{}
  \date{}
  \predate{}\postdate{}

@@

Marine reserves are implemented to achieve a variety of objectives, but
are seldom rigorously evaluated to determine whether those objectives
are met. In the rare cases when evaluations do take place, they
typically focus on ecological indicators and ignore other relevant
objectives such as socioeconomics and governance. And regardless of the
objectives, the diversity of locations, monitoring protocols, and
analysis approaches hinder the ability to compare results across case
studies. Moreover, analysis and evaluation of reserves is generally
conducted by outside researchers, not the reserve managers or users,
plausibly thereby hindering effective local management and rapid
response to change. We present a framework and tool, called ``MAREA'',
to overcome these challenges. Its purpose is to evaluate the extent to
which any given reserve has achieved its stated objectives. MAREA
provides specific guidance on data collection and formatting, and then
conducts rigorous causal inference analysis based on data input by the
user, providing real-time outputs about the effectiveness of the
reserve. MAREA's ease of use, standardization of state-of-the-art
inference methods, and ability to analyze marine reserve effectiveness
across ecological, socioeconomic, and governance objectives could
dramatically further our understanding and support of effective marine
reserve management.

@@

Empirical evidence shows that MPAs increase biomass {[}4,7{]}, enhance
resilience to climatic impacts {[}8,9{]}, and preserve genetic diversity
{[}10{]}. Compared to MPAs that grant partial protection, marine
reserves have higher levels of biomass, density, richness, and larger
organisms {[}3,11--13{]}. However, these effects are often measured as
biological changes within the reserves through time, and many lack a
control site for comparison {[}14{]}. This approach does not account for
other factors (\emph{e.g.} system-level changes in productivity caused
by predatory release {[}15{]}; or favorable environmental conditions
{[}16{]}) for which one must control {[}17{]} in order to causally
attribute a biological change to the reserve. Other studies have used a
control-impact comparison approach that uses control sites but does not
address temporal variability {[}4,7,18--20{]}.

A smaller fraction of studies have used a before-after-control-impact
(\emph{i.e.} BACI) design comparing reserves to control sites before and
after implementation {[}4,21,22{]}, which allows the use of causal
inference techniques that estimate the effect of the reserve. For
example, in ref {[}21{]} authors use a BACI design and observe increases
in lobster catches --a proxy for abundances-- after reserve
implementation for protected and control sites. However, the temporal
changes in the reserve were greater than in the control site, suggesting
a positive effect of the reserve on lobster catches. But even when
proper causal inference can be drawn, results are often different across
reserves. Effects of reserves on ecological and economic outcomes are
highly heterogeneous, and often depend on the specific ecological,
economic, and social context.

@@

An increasingly popular way to make science accessible, reproducible,
scalable, and replicable is through Open Science and the development of
open-access tools {[}28{]}. The Ocean Health Index {[}29,30{]}, for
example, successfully standardized a way to measure the health and
benefits of the oceans. This approach has been implemented at a global
scale, but also at country-level {[}31{]}, and regionally {[}32,33{]}.
Open access tools are not limited to conservation, and have also been
developed to evaluate fishery performance {[}34,35{]}, design
territorial use rights for fisheries {[}36{]}, and improve decision
making in the hydro power industry {[}37{]}.

The purpose of this paper is to describe a user-friendly tool, called
``MAREA'', to rigorously systematize the evaluation of marine reserve
effectiveness in terms of fisheries and marine conservation goals. The
tool is in the form of an open-source application that uses
state-of-the-art methods from program evaluation to compare a reserve to
control sites along a number of biological, economic, and governance
dimensions. We first provide a list of commonly stated management
objectives and match them to appropriate indicators. We then develop a
simple approach to analyzing these indicators building on causal
inference techniques {[}21{]}, which help us understand the effect of
management interventions {[}25,38{]}. To implement the analytical
approach, we introduce the Marine Reserve Evaluation Application
(MAREA), an open source, web--based tool that automates the framework
described in this paper and enables its broader use. Finally, we present
a case study on the evaluation of a marine reserve established by the
fishers of Isla Natividad (Mexico) in 2006, to demonstrate the potential
of MAREA.

@@

Throughout this study, we will refer to the stated goals for which a
marine reserve was designed as ``objectives.'' This work was motivated
by the case of Mexico, where 39 reserves have been implemented over the
past five years to achieve objectives such as increasing productivity in
nearby waters or recovery of overexploited species; most of these
reserves have never been formally evaluated for effectiveness at meeting
those objectives. Thus, our focus was on identifying common objectives
of marine reserves in Mexico. However, a literature review and
discussions with marine reserve researchers suggested that the
objectives driving Mexican marine reserve implementation are similar to
those in the rest of the world. Thus, we group these objectives into
seven major categories that may be applied to marine reserves worldwide.
Any given reserve may have been implemented to meet one or more of
these. The list includes objectives stated in legislation {[}39,40{]}
and official documents such as the Technical Justification Studies
(\emph{Estudios Técnicos Justificativos}), agreements, and decrees
associated to these areas:

@@

Based on these seven objectives, we determined a set of associated
indicators to evaluate reserve effectiveness. These indicators are
specific variables on which data could be collected and analyzed, to
ultimately determine whether the corresponding objective was causally
being achieved by the marine reserve. The list of indicators was
compiled through a review of scientific literature in which we
identified indicators that were used to measure similar
objectives{[}3--5,7,11,13,14,18--21,23,24,41--44{]}. A first filter
eliminated indicators for which baseline data do not typically exist in
Mexico. The preliminary list of indicators was reviewed at a workshop
with participation of members from Mexican fishery management agencies
and non-government organizations. Later, these were presented to fishers
from the Ensenada Fishing Cooperative (\emph{S.C.P.P. Ensenada}), in El
Rosario, Baja California, who provided input. Our final list of
indicators includes those identified in review works {[}4,44{]}.

Indicators are divided into three main categories: biological,
socioeconomic, and governance (Table 1). The nine biological indicators
focus on fish and invertebrate communities that are evaluated using
underwater ecological surveys performed inside and outside the reserve
(see Data and Analysis section for specific sampling design and
methodologies). Five socioeconomic indicators reflect the performance of
the fishery in terms of landings, income from landings, and availability
of alternative livelihoods. Fifteen governance indicators describe the
governance structures under which the community operates (\emph{e.g.},
access rights to the fishery, number of fishers, legal recognition of
the reserve). Most biological and socioeconomic indicators are
quantitative and require a numerical entry (\emph{e.g.} Fish biomass)
while all governance indicators, one biological indicator, and one
socioeconomic indicator are qualitative and rely on a descriptive entry
(\emph{e.g.} Reasoning for reserve location). Many of them specifically
measure an outcome of the reserve, though some are designed to further
the understanding of the mechanisms driving a reserve's performance. In
that sense, most biological and socioeconomic indicators are outcome
variables. On the other hand, governance indicators are viewed as
possible explanatory variables of reserve performance. Whenever an
indicator is applied to ``Target species'', it means that the indicator
can be used for all species (\emph{e.g.} Fish Biomass) and/or for
individual species that are either the conservation target of the
reserve or are of particular economic or ecological interest
(\emph{e.g.} Grouper Biomass). Finally, indicators B3 and B4 are
different in that B3 only looks at the density of organisms above size
at first maturity (related to reproductive potential), while B4 measures
the density of all fish or of a target species. Each indicator targets
different plausible desired outcomes, like increased reproductive
potential (\emph{i.e.} B3; {[}45{]}) or having more fish -regardless of
their size- to attract tourism (\emph{i.e.} B4). Table 1 presents the
proposed indicators, and Table 2 shows how objectives are matched with
biological and socioeconomic indicators. Governance indicators are
excluded from Table 2, but should be considered for every objective as
each serves as a plausible explanatory variable for reserve performance.

@@

\textbf{Table 1. List of indicators to evaluate the effectiveness of
no-take marine reserves.}

\begin{table}[H]
\centering
\resizebox{\linewidth}{!}{\begin{tabular}{l|l|l|l}
\hline
\bfseries{Code} & \bfseries{Indicator} & \bfseries{Data type} & \bfseries{Unit}\\
\hline
\multicolumn{4}{l}{\textbf{Biological}}\\
\hline
\hspace{1em}B1 & Shannon diversity index & Continuous & \\
\hline
\hspace{1em}B2 & Species richness & Discrete & Number of species/transect\\
\hline
\hspace{1em}B3 & Density of mature organisms & Continuous & Percent\\
\hline
\hspace{1em}B4 & Density* & Continuous & Organisms/transect\\
\hline
\hspace{1em}B5 & Natural Disturbance & Descriptive & \\
\hline
\hspace{1em}B6 & Mean Trophic Level & Continuous & \\
\hline
\hspace{1em}B7 & Biomass* & Continuous & kg/transect\\
\hline
\multicolumn{4}{l}{\textbf{Socioeconomic}}\\
\hline
\hspace{1em}S1 & Total landings* & Continuous & kg\\
\hline
\hspace{1em}S2 & Income from total landings* & Continuous & \$\\
\hline
\hspace{1em}S3 & Alternative economic opportunities & Ordinal & \\
\hline
\multicolumn{4}{l}{\textbf{Governance}}\\
\hline
\hspace{1em}G1 & Access to the fishery & Categorical & \\
\hline
\hspace{1em}G2 & Number of fishers & Discrete & \\
\hline
\hspace{1em}G3 & Legal recognition of reserve & Binary & \\
\hline
\hspace{1em}G4 & Reserve type & Descriptive & \\
\hline
\hspace{1em}G5 & Illegal harvesting & Ordinal & \\
\hline
\hspace{1em}G6 & Management plan & Binary & \\
\hline
\hspace{1em}G7 & Reserve enforcement & Descriptive & \\
\hline
\hspace{1em}G8 & Size of reserve & Discrete & \\
\hline
\hspace{1em}G9 & Reasoning for reserve location & Descriptive & \\
\hline
\hspace{1em}G10 & Membership to fisher organizations & Binary & \\
\hline
\hspace{1em}G11 & Type of fisheries organizations & Categorical & \\
\hline
\hspace{1em}G12 & Representation & Ordinal & \\
\hline
\hspace{1em}G13 & Internal Regulation & Binary & \\
\hline
\hspace{1em}G14 & Perceived Effectiveness & Categorical & \\
\hline
\hspace{1em}G15 & Social Impact of Reserve & Categorical & \\
\hline
\end{tabular}}
\end{table}

* Indicates the indicator is applied to target species

\textbf{Table 2. Management objectives and respective performance
indicators.}

\begin{table}[H]
\centering
\resizebox{\linewidth}{!}{\begin{tabular}{>{\raggedright\arraybackslash}p{4cm}|l|l|l|l|l|l|l|l|l|l|l|l|l|l}
\hline
\bfseries{Objective} & \bfseries{B1} & \bfseries{B2} & \bfseries{B3} & \bfseries{B4} & \bfseries{B4*} & \bfseries{B5} & \bfseries{B6} & \bfseries{B7} & \bfseries{B7*} & \bfseries{S1} & \bfseries{S1*} & \bfseries{S2} & \bfseries{S2*} & \bfseries{S3}\\
\hline
Avoid overexploitation &  &  & x & x & x & x & x & x & x & x & x & x & x & x\\
\hline
Conserve species under a special protection &  &  & x &  & x & x &  &  & x & x &  & x &  & x\\
\hline
Maintain biological process & x & x &  & x &  & x & x & x &  &  &  &  &  & x\\
\hline
Improve fishery production in nearby waters &  &  &  & x & x & x &  & x & x & x & x & x & x & x\\
\hline
Preserve biological diversity and the ecosystem & x & x &  & x &  & x & x & x &  &  &  &  &  & x\\
\hline
Recover overexploited species &  &  & x &  & x & x &  &  & x &  & x &  & x & x\\
\hline
Recover species of economic interest &  &  & x &  & x & x &  &  & x &  & x &  & x & x\\
\hline
\end{tabular}}
\end{table}

Governance indicators are excluded from the table, but all should be
used for any objective.

* Indicates the indicator is applied to target species

@@

Socioeconomic data are often collected by fishers, natural resource
management agencies, or Civil Society Organizations (CSOs) by recording
landings, income, and sometimes prices for each species. To control for
inflation, income is adjusted with the country's consumer price index
{[}48{]}:

@@

While this model does not formally allow for causal inference, we can
still measure changes in mean landings and income before and after the
implementation of the reserve and provide valuable input. For both
models (eq. 1 and eq. 3), we estimate the model coefficients with
ordinary least squares, and calculate heteroskedastic--robust standard
errors {[}49{]}.

@@

We developed MAREA in R version 3.4.2 and R Studio 1.1.383 {[}51{]}
using the Shiny package {[}52{]}, to build an interactive web
application hosted on an open server; the MAREA app can be accessed at
\href{turfeffect.shinyapps.io/marea/}{turfeffect.shinyapps.io/marea}.
While the original version was developed in Spanish because it was aimed
for Mexico and other Latin-American countries, all of its content can be
translated by a translation widget available within the app.

MAREA is designed as a 6-step process, divided in tabs that appear upon
launching the app. The first tab introduces the app and summarizes the
evaluation process. Then, the user selects management objectives, which
MAREA automatically matches to appropriate indicators, based on Table 2.
Users can also select and deselect indicators based on their interests
and data availability by ``clicking'' on the check-boxes in MAREA. The
user can then load data on one or more reserves, using standard *.csv
text files; sample datasets are provided within MAREA. Once data have
been loaded, MAREA identifies all reserves in the data, and lets the
user select the reserve to be evaluated. At this point, the user can
also specify the year of implementation of the reserve, reserve
dimensions, and indicate target species that are of particular
management interest. MAREA provides the user with a section to confirm
that all the decisions made leading up to that point are correct. Once
the user has confirmed all input data, objectives, and other
information, MAREA performs the formal program evaluation analyses
discussed above. For a typical data set, the automated analysis step
takes less than one second. Finally, the user is taken to the results
tab where all results are presented in a simple format. The user can
also download a more comprehensive technical report produced in *.pdf
format.

@@

\textbf{Fig 3. General location of Isla Natividad (left) and map of the
island (right).} The marine reserve polygon is indicated in red, and the
approximate location of control sites is indicated by blue squares (B =
Babencho, D = La Dulce). Shapefiles for Mexican coastline and the United
States were obtained from the Instituto Nacional de Estadistica,
Geografia e Informatica of Mexico (INEGI:
www.inegi.org.mx/geo/contenidos/geoestadistica/m\_geoestadistico.aspx)
and the tmap R package {[}58{]}, respectively.

In 2006, the Isla Natividad community established a biological baseline
following the data collection protocol described in this study. The
community then implemented two community-based marine reserves within
their TURF {[}8,41,59{]} after establishing a baseline for the
soon-to-be reserves and control sites. Evidence suggest that these
reserves have been effective at enhancing resilience to climate
variations {[}8{]} and preserving genetic diversity of high value
commercial species such as abalone {[}10{]}. These ecological benefits
have been translated into economic benefits, enhancing population
persistence and bolstering abalone fisheries {[}43{]}. For the purpose
of this evaluation, we focused on the ``La Plana / Las Cuevas'' marine
reserve, located at the southern end of the island (Fig. 3) and its
corresponding control site ``La Dulce / Babencho''.

@@

\textbf{Table 3. Summary of average treatment effect of the reserve on
biological indicators.}
